### 2020-09-28 doc/manuscript/coauthor_feedback_to_incorporate/Paolo_feedback_2020_09.RmD

Dear Kristina,

thank you for including me in this interesting study and for sending over the draft.

Here some thoughts:

1) In the intro you state that tree rings can be used to estimate forest woody productivity. this is certainly true but a thorough discussion of the sampling bias (Cherubini et al. For Ecol manage 1998) and necessary sampling design (Nehrbass-Ahles Gl Ch BIol 2014) should be included.

2) tree size affects tree growth rates, true, but is it really the diameter or rather tree age, tree height, crown size, root depth? A physiological discussion is probably needed here.

3) How were the sites selected? Site selection is crucial to correctly interpret the results.

4) The diversity of growth trends in relation to DBH is largely attributable to species ecology and stand history, true; stand dynamics strongly affect stand structure and therefore tree diameter and growth trends.

5) “small stems that will never enter the cohort of trees sampled by coring”, see previous comment on sampling bias (Cherubini et al. 1998)

6) “deposition of sulfur dioxide”: is it still a problem of relecance?

7) “the impact of rising CO2 on long-term growth trends reach inconsistent conclusions”: because of the influence of stand dynamics.

8) The ITRDB doesn’t contain any structure for storing DBH records, but ring widths allow the reconstruction of diameter over time.

Hope it helps.

Best regards,

Paolo
--