### 2021-03-09 doc/manuscript/tables_figures/tables.Rmd
started hypothesis table
#119

---
title: " "

output: 
  pdf_document:
    fig_caption: yes

header-includes:  \usepackage{float}
                  \usepackage{caption}
                  \captionsetup[table]{font=footnotesize}
                  \captionsetup[figure]{font=footnotesize}
                  \captionsetup[figure]{labelformat=empty}
                  \captionsetup[table]{labelformat=empty}
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
options(knitr.table.format = "latex")
```

\newpage

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
sites <- read.csv("sites.csv", stringsAsFactors = FALSE, check.names = FALSE)
sites_table <- subset(sites, select = -c(MAT, JulyTsd, JanTsd, MAPsd) ) #drop MAT column, plus columns with sd
kable(sites_table, format = "latex", booktabs = TRUE,  escape = F)%>%
  add_header_above(c(" ", " "," ", "1950 - 2019 climate" = 3, " " , " "," "," ")) %>%
  column_spec(1, width = "1cm") %>%
  column_spec(2, width = "3cm") %>%
  column_spec(3, width = "3cm") %>%
  column_spec(4:6, width = "1cm") %>%
  #column_spec(6, width = ".5cm") %>%
  column_spec(7, width = "1.4cm") %>%
  column_spec(8, width = "1cm") %>%
  #column_spec(9, width = "2cm") %>%
  column_spec(10, width = "4cm") %>%
  kable_styling(latex_options = c("scale_down", "hold_position"), protect_latex = T) 
```

\newpage

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
hypotheses <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
hypotheses_table <- subset(hypotheses, select = -c(category, percent.sites, percent.species.site.combinations) ) #drop MAT column, plus columns with sd
kable(hypotheses_table, format = "latex", booktabs = TRUE,  escape = F)%>%

  column_spec(1, width = "12cm") %>%
  column_spec(2, width = "5cm") %>%

  pack_rows("Annual growth is jointly influenced by water, temperature, DBH, and time. ", 1, 2, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
  pack_rows("Drought limits tree growth, but the response to water is nonlinear.", 3, 5, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
  pack_rows("Temperature has predominantly negative, nonlinear effects on growth.", 6, 9, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
  pack_rows("DBH", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%

  kable_styling(latex_options = c( "hold_position", "repeat header"), protect_latex = T, font_size = 7)

```